---
output: html_document
title: "Group-Project"
date: "2024-04-17"
---
<style>
body {
    background-color: black;
    color: white;
}
</style>
---


```{r setup, include=FALSE}
#Set your working directory
setwd("/Users/amiraparker/Desktop/Stats Class")

#Open packages 

library(readxl)
library(readr)
library(psych)
library(car)
library(dplyr)
library(ggpubr)
library(jtools)
library(effectsize)

#load data set
project_data <- read_csv("Group 1 - ANOVA 7x4 Data.csv")
project_data <- as.data.frame(project_data)

attach(project_data)

project_data

project_data <- project_data[, -which(names(project_data) == "BDI")]

knitr::opts_chunk$set(echo = TRUE)
```



## <span style="color:orange"> Step 1: Examine Data </span>

### Create a table that indicates if we have "balanced" ANOVA
Each cell contains the same number of participants across our factors.
```{r Ensuring We have a Blanced ANOVA}

tbl_count_project <- table(Race_Ethnicity,Employment)
tbl_count_project


```
We can see in the output that there are 5 individuals in each possible cell, which means we DO have a balanced design and can run a standard Two-Way ANOVA.

### Look at the proportion of individuals in each cell.
```{r Proportion of individuals in each cell}
tbl_percent_project <- round(prop.table(tbl_count_project), digits=2)
tbl_percent_project
```

### Add the sums for each factor level
```{r Add sums for each factor level}
tbl_cross_project <- table(Race_Ethnicity,Employment)
addmargins(tbl_cross_project)
```

### Display summed table in Bar Chart
```{r Bar Chart, fig.width=15, message=FALSE, warning=FALSE, echo=FALSE}
barplot(tbl_cross_project, xlab='Employment',ylab='Frequency',main="Frequency Employment by Race",
        col=c("darkblue","lightcyan","turquoise")
        ,legend=rownames(tbl_cross_project), args.legend = list(x = "topleft"))

```

### Examine variablity across factor levels using box/scatter plots
```{r Variability across factor levels, fig.width=15, message=FALSE, warning=FALSE, echo=FALSE}
ggboxplot(project_data, x = "Race_Ethnicity", y = "ASQ", color = "Employment",
          palette = c("#00AFBB", "#E7B800","#B7B800","#F7B800"))
ggline(project_data, x = "Race_Ethnicity", y = "ASQ", color = "Employment",
       add = c("mean_se", "dotplot"),
       palette = c("#00AFBB", "#E7B800","#B7B800","#F7B800"))


ggboxplot(project_data, x = "Employment", y = "ASQ", color = "Race_Ethnicity",
          palette = c("#00AFBB", "#E7B800", "#228C22", "#618C22", 
                      "#10AFBB", "#89B800", "#988C22"))
ggline(project_data, x = "Employment", y = "ASQ", color = "Race_Ethnicity",
       add = c("mean_se", "dotplot"),
       palette = c("#00AFBB", "#E7B800", "#228C22", "#618C22", 
                   "#10AFBB", "#89B800", "#988C22"))
```

### <span style="color:blue"> Mean comparison model for race_ethnicity </span>
Scatter plot graph with the means and standard error bars.
```{r Race_Ethnicity Mean Comparison, fig.width=15, message=FALSE, warning=FALSE, echo=FALSE}
ggline(project_data, x = "Race_Ethnicity", y = "ASQ", 
       add = c("mean_se", "jitter"), 
       order = c("American_Indian_Alaskan", "Asian", "Black_AA", 
                 "Hispanic_Latino", "Multi-Racial", "Nat_Hawaiian_PI","White"),
       ylab = "ASQ", xlab = "Race_Ethnicity")
```

### <span style="color:purple"> Mean comparison model for employment </span>
Scatter plot graph with the means and standard error bars.
```{r Employment Mean Comparison, warning=FALSE, fig.width=15, message=FALSE, echo=FALSE}
ggline(project_data, x = "Employment", y = "ASQ", 
       add = c("mean_se", "jitter"), 
       order = c("Unemployed","Self_Employed","Employed_PT","Employed_FT"),
       ylab = "ASQ", xlab = "Employment")
```



# <span style="color:orange"> Step 2: Running the 2-way ANOVA </span>

```{r 2-way Anova}
project_2way <- aov(ASQ ~ Race_Ethnicity + Employment + Race_Ethnicity:Employment, data = project_data)

summary(project_2way)

```

# <span style="color:orange"> Step 3: Testing Simple Main Effects for Our Factors
Using one way anovas, we test simple main effect of race_ethnicity at all levels of our Employment factor("Unemployed","Self_Employed","Employed_PT","Employed_FT") </span>


## <span style="color:blue">  Simple Main Effect for Race_Ethnicity </span>


### <span style="color:#3399FF"> Simple Main Effect for Race_Ethnicity for Unemployed-only level </span>
```{r Race/Ethnicity Main Effect - Unemployed}
unemployed_data <- subset(project_data, Employment == "Unemployed")
unemployed_anova <- aov(ASQ ~ Race_Ethnicity, data = unemployed_data)
summary(unemployed_anova)
TukeyHSD(unemployed_anova)
```
**The Race_Ethnicity simple main effect was significant indicating that at least twomeans differ across the Race_Ethnicity groups within the unemployed employment level. since we have 7 levels in Race_Ethnicity, we  ran a post-hoc comparison and found the following pairs with sgnificant differences: Asian-American_Indian_Alaskan, White-American_Indian_Alaskan, Black_AA-Asian,Multi-Racial-Asian, Nat_Hawaiian_PI-Asian, Hispanic_Latino-Black_AA, White-Black_AA, White-Hispanic_Latino, White-Multi-Racial, White-Nat_Hawaiian_PI**

### <span style="color:#3399FF"> Simple Main Effect for Race_Ethnicity for Selfemployed-only level </span>
```{r Race/Ethnicity Main Effect - Self_Employed}
selfemployed_data <- subset(project_data, Employment == "Self_Employed")
selfemployed_anova <- aov(ASQ ~ Race_Ethnicity, data = selfemployed_data)
summary(selfemployed_anova)
TukeyHSD(selfemployed_anova)
```
**The Race_Ethnicity simple main effect was significant indicating that at least two means differ across the Race_Ethnicity groups within the self-employed employment level. 
We ran a post-hoc comparison and found the following pairs with significant differences: White-American_Indian_Alaskan, White-Black_AA**

### <span style="color:#3399FF"> Simple Main Effect for Race_Ethnicity for Part Time employed-only level </span>
```{r Race/Ethnicity Main Effect - Employed_PT}
employedpt_data <- subset(project_data, Employment == "Employed_PT")
employedpt_anova <- aov(ASQ ~ Race_Ethnicity, data = employedpt_data)
summary(employedpt_anova)
TukeyHSD(employedpt_anova)
```
**The Race_Ethnicity simple main effect was significant indicating that at least two means differ across the Race_Ethnicity groups within the parttime-employed employment level. 
We ran a post-hoc comparison and found the following pairs with significant differences: White-American_Indian_Alaskan, White-Black_AA, White-Multi-Racial**

### <span style="color:#3399FF"> Simple Main Effect for Race_Ethnicity for Full Time employed-only level </span>
```{r Race/Ethnicity Main Effect - Employed_FT}
employedft_data <- subset(project_data, Employment == "Employed_FT")
employedft_anova <- aov(ASQ ~ Race_Ethnicity, data = employedft_data)
summary(employedft_anova)
TukeyHSD(employedft_anova)
```
**The Race_Ethnicity simple main effect was significant indicating that at least two means differ across the Race_Ethnicity groups within the fulltime-employed employment level. 
We ran a post-hoc comparison and found the following pairs with significant differences: White-American_Indian_Alaskan, White-Black_AA, White-Multi-Racial, White-Nat_Hawaiian_PI**

### <span style="color:#3399FF"> Simple Main Effect for Race_Ethnicity Visual </span>
```{r Visualizing Simple Main Effect for Race_Ethnicity, echo=FALSE, fig.width=15, message=FALSE, warning=FALSE}
ggline(project_data, x = "Employment", y = "ASQ", color = "Race_Ethnicity",
       add = c("mean_se", "dotplot"),
       palette = c("#00AFBB", "#E7B800", "#228C22", "#618C22", 
                   "#10AFBB", "#89B800", "#988C22"))
```


## <span style="color:purple"> Simple Main Effect for Employment </span>
Using one way anovas, we test simple main effect of race_ethnicity at all 7 levels of our Race_ethnicity factor

### <span style="color:#9370DB"> Simple Main Effect for Employment for American_Indian_Alaskan-only level </span>
```{r Employment Main Effect - American_Indian_Alaskan}
AmInAl_data <- subset(project_data, Race_Ethnicity == "American_Indian_Alaskan")
AmInAl_anova <- aov(ASQ ~ Employment, data = AmInAl_data)
summary(AmInAl_anova)
TukeyHSD(AmInAl_anova)
```
**It seems from this output, these are pairs with sgnificant differences: 
Unemployed-Employed_PT, Unemployed-Self_Employed. From this output, we have a significant simple main effect, where there is a significant difference between ASQ scores for Unemployed, parttime employed and self employed means for those who arewho are American_Indian_Alaskan, where those who are unemployed have lower ASQ than part time employed and self-time employed**

### <span style="color:#9370DB"> Simple Main Effect for Employment for Asian-only leve </span>
```{r Employment Main Effect - Asian}
asian_data <- subset(project_data, Race_Ethnicity == "Asian")
asian_anova <- aov(ASQ ~ Employment, data = asian_data)
summary(asian_anova)
TukeyHSD(asian_anova)
```
**It seems from this output, these are pairs with sgnificant differences: 
#Unemployed-Employed_PT, Unemployed-Self_Employed. From this output, we have a significant simple main effect, where there is a significant difference between ASQ scores for Unemployed, parttime employed and self employed means for those who are who are Asian where those who are unemployed have lower ASQ than part time employed and self-time employed**

### <span style="color:#9370DB"> Simple Main Effect for Employment for Black-only level </span>
```{r Employment Main Effect - Black}
black_data <- subset(project_data, Race_Ethnicity == "Black_AA")
black_anova <- aov(ASQ ~ Employment, data = black_data)
summary(black_anova)
TukeyHSD(black_anova)
```
**It seems from this output, these are pairs with sgnificant differences: Unemployed-Self_Employed. From this output, we have a significant simple main effect, where there is a significant difference between ASQ scores for Unemployed, self employed means for those who are who are Black where those who are unemployed have lower ASQ than those who are self employed**

### <span style="color:#9370DB"> Simple Main Effect for Employment for Hispanic-only level </span>
```{r Employment Main Effect - Hispanic_Latino}
hisp_data <- subset(project_data, Race_Ethnicity == "Hispanic_Latino")
hisp_anova <- aov(ASQ ~ Employment, data = hisp_data)
summary(hisp_anova)
TukeyHSD(hisp_anova)
```
**It seems from this output, these are pairs with sgnificant differences:  Unemployed-Self_Employed. From this output, we have a significant simple main effect, where there is  a significant difference between ASQ scores for Unemployed, self employed means for those who arewho are Hispanic where those who are unemployed have lower ASQ than those who are self employed**

### <span style="color:#9370DB"> Simple Main Effect for Employment for Multi-Race-only level </span>
```{r Employment Main Effect - Multirace}
multi_data <- subset(project_data, Race_Ethnicity == "Multi-Racial")
multi_anova <- aov(ASQ ~ Employment, data = multi_data)
summary(multi_anova)
TukeyHSD(multi_anova)
```
**It seems from this output, these are pairs with sgnificant differences: 
Unemployed-Employed_PT, Unemployed-Self_Employed. From this output, we have a significant simple main effect, where there is a significant difference between ASQ scores for Unemployed, parttime employed and self employed means for those who are
who are Multi-race where those who are unemployed have lower ASQ than part time employed and self-time employed**

### <span style="color:#9370DB"> Simple Main Effect for Employment for Nat_Hawaiian_PI-only level </span>
```{r Employment Main Effect - Nat_Hawaiian_PI}
NatHawPI_data <- subset(project_data, Race_Ethnicity == "Nat_Hawaiian_PI")
NatHawPI_anova <- aov(ASQ ~ Employment, data = NatHawPI_data)
summary(NatHawPI_anova)
TukeyHSD(NatHawPI_anova)
```
**It seems from this output, these are pairs with sgnificant differences: 
Unemployed-Self_Employed From this output, we have a significant simple main effect, where there is a significant difference between ASQ scores for Unemployed, and self employed means for those who are who are Native Hawaiian Pacific Islander where those who are unemployed have lower ASQ than self employed** 

### <span style="color:#9370DB"> Simple Main Effect for Employment for White-only level </span>
```{r Employment Main Effect - White}
white_data <- subset(project_data, Race_Ethnicity == "White")
white_anova <- aov(ASQ ~ Employment, data = white_data)
summary(white_anova)
TukeyHSD(white_anova)
```
**It seems from this output, these are pairs with significant differences: Unemployed-Self_Employed From this output, we have a significant simple main effect, where there is  a significant difference between ASQ scores for Unemployed, and self employed means for those who are who are White where those who are unemployed have lower ASQ than self employed**

### <span style="color:#9370DB"> Simple Main Effect for Employment Visual </span>
```{r Visualizing Simple Main Effect for Employment, echo=FALSE, fig.width=15, message=FALSE, warning=FALSE}
ggline(project_data, x = "Race_Ethnicity", y = "ASQ", color = "Employment",
       add = c("mean_se", "dotplot"),
       palette = c("#00AFBB", "#E7B800", "#228C22", "#618C22"))
```

## <span style="color:orange"> Step 4: Getting Descriptive Statistics </span>
The following includes descriptive statistics needed for reporting results and/or creating a means table.

### <span style="color:blue"> Race_Ethnicity Descriptives </span>
```{r Race_Ethnicity Descriptives, echo=FALSE}

table_race <- as.data.frame(group_by(project_data, Race_Ethnicity) %>%
                             summarise(
                               count = n(),
                               mean = mean(ASQ, na.rm = TRUE),
                               sd = sd(ASQ, na.rm = TRUE)
                             ))
table_race
```
### <span style="color:purple"> Employment Descriptives </span>
```{r Employment Descriptives, echo=FALSE}

table_employment <- as.data.frame(group_by(project_data, Employment) %>%
                              summarise(
                                count = n(),
                                mean = mean(ASQ, na.rm = TRUE),
                                sd = sd(ASQ, na.rm = TRUE)
                              ))
table_employment
```
### Means Table 
```{r Mean Table, echo=FALSE, message=FALSE, warning=FALSE}
table_all_project <- as.data.frame(group_by(project_data, Race_Ethnicity, Employment) %>%
                summarise(
                   count = n(),
                    mean = mean(ASQ, na.rm = TRUE),
                    sd = sd(ASQ, na.rm = TRUE)
                    ))
table_all_project
```

### Grand Mean for ANOVA model
```{r Grand Mean}

describe(ASQ)
```

## <span style="color:orange"> Step 5: Partial Eta-Squared  </span>
Calculate the partial eta-squared values that align to the various effects within our GLM 2-way ANOVA model.

### Ensure we are specifying the correct model
```{r partial eta-squared - specify correct model}

m_2way_project <- aov(ASQ ~ Race_Ethnicity + Employment + Race_Ethnicity:Employment, data = project_data)
```

### Run the partial eta-squared estimates for our 2-way model effects. 
```{r run partial eta-squared, message=FALSE, warning=FALSE}
eta_squared(car::Anova(m_2way_project, partial = TRUE, type = 2))
```
We can see from these results that interaction Race_Ethnicity:Employment has a small effect size (.02). Employment and Race_ethnicity have large effect sizes: (.44) and (.53) respectively. 

## Testing Assumptions

### Homogeneity of Variance
```{r Homogeneity of Variance}
leveneTest(ASQ ~ Race_Ethnicity*Employment, data = project_data)
```
The assumption is maintained here since the p-value of Levene's test was > .05.



